Gaia's constraints on the extremely massive ($\sim$$150\,M_\odot$) runaway star candidate VFTS682

### 2018-06-10 VFTS682_APJL.tex
proof read sec.2 and filled tab 1

How massive stars form is one of the major longstanding questions in astrophysics
\citep[e.g.,][]{zinnecker:07}. Improving our understanding of massive star formation, and its
possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies, but also for understanding the transients that mark their death and the compact remnants they leave behind.
%
Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
% \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
evolve fast, typically reside in dense groups and remain enshrouded in
their parent cloud during the entirety of their formation
process. Important progress has been made on the theoretical side,
\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally very expensive and therefore remain challenging.

@@

VFTS682 stands out by its relative isolation at a projected distance of 119.4 arcseconds, corresponding to 
$\sim$$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
considered two possible explanation for the offset: either
the star formed in situ as an isolated massive star, % , but remark that this
                                % poses a challenge for starformation
                                % theories (They don't)
or it was ejected from  R136. N-body simulations % of dynamical interaction in star clusters
indicate that the ejection of very massive stars like VFTS682 is expected \citep[e.g.][]{fujii:11, banerjee:12}.  This is supported by the recent findings of other massive runaway stars in the region based on proper motion studies.   \citet{platais:15,platais:18} analyze multi-epoch \emph{Hubble Space Telescope} (HST) photometry and identify 10 stars, which appear to be ejected from R136.   \citet{lennon:18} investigate the kinematics of  isolated O-type stars in the region using the second \emph{Gaia} data release \cite[DR2,][]{gaia:16,brown:18} and show that the proper motion, postion and direction of the $\sim100\Msun$ star VFTS16 is consistent with a runaway origin from R136.

@@

       \hline
       \multicolumn{5}{l}{\emph{Selected stellar parameters VFTS 682}} \\
     & present day mass  & $[M_\odot]$ & $137.8^{+27.5}_
                                           {-15.9}$ &
                                                    \\
      & initial mass& $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ &
      \\
      &age & [Myr] & $1.0\pm0.2$ & \\
      &mass loss rate &  & $\SdM{xxx}$ & \\
      &helium abundance &  & $\SdM{xxx}$ & \\
      \hline

    \end{tabular}
    \tablecomments
    { 
    }
  \end{center}
  \label{tab:star}
\end{table}

@@

       \hline
           \multicolumn{5}{l}{\emph{Gaia's proper motion for VFTS 682 and average of the region}} \\
          &$\mu_\mathrm{RA}$&[$\masyr$] & $1.843\pm 0.070$
                     & (1) \\        
               &$\mu_\mathrm{DEC}$&[$\masyr$] & $0.786\pm 0.080$
                     &  (1) \\        
                 & Correlation  ($\mu_\mathrm{RA}$, $\mu_\mathrm{DEC}$) &[$\masyr$] & $0.023$
                        & (1)  \\         
     %  \multicolumn{5}{l}{\emph{Gaia's average proper motion in the region}} \\         
       &$\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$&[$\masyr$] & $1.74\pm0.01$
                        & (2) \\
      &$\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$&[$\masyr$]
                & $0.70\pm0.02$ &  (2)\\
\hline

@@

    %  &RA &[degree] &  \phantom{-}84.73 % $\pm$  0.03
   %                     & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
    %  &DEC & [degree] & -69.07 % $\pm$  0.05
     %                   & \\[5pt]
      \multicolumn{5}{l}{\emph{Gaia's relative proper motion for VFTS 682 }} \\                 
      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & 0.103$\pm$0.080 & (1,4) \\
      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & 0.086$\pm$0.100 &  (1,4) \\
      &$\delta\mu_\mathrm{}$  &[$\masyr$] & \SdM{0.13}$\pm$\SdM{0.09} &  (1,4) \\
                 &$v_\mathrm{2d}$  &[$\kms$] & \SdM{32}$\pm$\SdM{21} & (1,4)\\  
                     &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
                         &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
 %

 \hline     
          \multicolumn{5}{l}{\emph{HST's independent determination of the relative proper motion for VFTS 682}} \\  
      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & 0.01$\pm$0.13 & (3) \\
      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & 0.20$\pm$0.10 &  (3) \\
       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & 0.20$\pm$0.10 &  (3) \\
                  &$v_\mathrm{2d, HST}$  &[$\kms$] & \SdM{$47\pm24$} & (3)\\  
                      &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
 %

@@

    \end{tabular}
    \tablecomments
    { Proper motions are based on Gaia data unless indicated otherwise in subscript. The angle $\theta$ is defined such that $\theta=0$ corresponds to the direction pointing away from R136.
    $^1$\cite{Gaia-Collaboration+2018}, 
    $^2$\cite{lennon:18}, 
    $^3$\cite{platais:18},  
    $^4${\color{blue}This work}
    %The error on the RA and DEC positions, are of order $\sim$$10^{-2}\masyr$ in \emph{Gaia} DR2. %REPETITION: %  We adopt the peculiar
      % radial velocity calculated using the  NV $\lambda4944$ line, which is
      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
    }
  \end{center}
  \label{tab:vfts682}
\end{table}

@@

The star WNh5 star VFTS682, located at right ascension (RA)
05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
%
%of spectral type WNh5 using multi-epoch spectra covering $\lambda$4000--7000 from the
%VFTS survey. % The spectra available covered $\lambda$4000--7000 and
% were taken at multiple epochs.
%
\citet{bestenlehner:11}  analyzed the spectra to infer the stellar parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of the brightest stars in the region. The star is unlikely to have a close companion, because of the absence of clear radial velocity variations \citet{bestenlehner:11}.  Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see Tab~ \label{tab:vfts682} for an overview of the parameters.

>>>>

How massive stars form is one of the major longstanding questions in astrophysics
\citep[e.g.,][]{zinnecker:07}. Improving our understanding of massive star formation, and its
possible dependence on environment and metallicity, is crucial for understanding the role massive stars play within their host galaxies, but also for understanding the transients that mark their death and the compact remnants they leave behind.
%
Obtaining clues from observations has been challenging, because massive stars are intrinsically rare, 
% \citep[e.g.,][]{salpeter:55,kroupa:01, schneider:18}, 
evolve fast, typically reside in dense groups, and remain enshrouded in
their parent cloud during the entirety of their formation
process. Important progress has been made on the theoretical side,
\citep[e.g.][]{bate:09,kuiper:15,rosen:16}, but the simulations of this multi-scale and multi-physics problem are computationally very expensive and therefore remain challenging.

@@

VFTS682 stands out by its relative isolation at a projected distance of 119.4 arcseconds, corresponding to 
$\sim$$29$\,pc, from  the star cluster R136. \citet{bestenlehner:11}
considered two possible explanation for the offset: either
the star formed in situ as an isolated massive star, % , but remark that this
                                % poses a challenge for starformation
                                % theories (They don't)
or it was ejected from  R136. N-body simulations % of dynamical interaction in star clusters
indicate that the ejection of very massive stars like VFTS682 is
expected \citep[e.g.][]{fujii:11, banerjee:12}.  This is supported by
the recent findings of other massive runaway stars in the region based
on proper motion studies.

\citet{platais:15,platais:18} analyzed multi-epoch \emph{Hubble Space
  Telescope} (HST) photometry and identified 10 stars likely ejected
from R136.   \citet{lennon:18} investigate the kinematics of  isolated
O-type stars in the region using the second \emph{Gaia} data release
\cite[DR2,][]{gaia:16,brown:18} and show that the proper motion,
postion and direction of the $\sim100\Msun$ star VFTS16 is consistent
with a runaway origin from R136. They also found a less clear case for VFTS72.

@@

       \hline
       \multicolumn{5}{l}{\emph{Selected stellar parameters VFTS 682}}
      \\
      \hline
     & present day mass  & $[M_\odot]$ & $137.8^{+27.5}_
                                           {-15.9}$ & (1)
                                                    \\
      & initial mass& $[M_\odot]$ & $150.0^{+28.7}_{-17.4}$ & (1)
      \\
      &age & [Myr] & $1.0\pm0.2$ & (1) \\
      &mass loss rate & [$M_\odot \ \mathrm{yr}^{-1}$] & $10^{-4.1\pm0.2}$ & (2)\\
      &\multirow{2}{*}{surface helium abundance} &  & 0.45 & (2)\\
      & &  & 0.49 & (3)\\
      \hline

    \end{tabular}
    \tablecomments
    { 
      (1)~\cite{schneider:18}
      (2)~\cite{bestenlehner:11}
      (3)~\cite{rubio-diez:17}
    }
  \end{center}
  \label{tab:star}
\end{table}

@@

       \hline
           \multicolumn{5}{l}{\emph{Gaia's proper motion for VFTS 682
      and average of the region}} \\
      \hline
          &$\mu_\mathrm{RA}$&[$\masyr$] & $1.843\pm 0.070$
                     & (1) \\        
               &$\mu_\mathrm{DEC}$&[$\masyr$] & $0.786\pm 0.080$
                     &  (1) \\        
                 & Correlation  ($\mu_\mathrm{RA}$, $\mu_\mathrm{DEC}$) &[$\masyr$] & $0.023$
                        & (1)  \\         
     %  \multicolumn{5}{l}{\emph{Gaia's average proper motion in the region}} \\         
       &$\langle\mu_\mathrm{RA}\rangle_\mathrm{R136}$&[$\masyr$] & $1.74\pm0.01$
                        & (2) \\
      &$\langle\mu_\mathrm{DEC}\rangle_\mathrm{R136}$&[$\masyr$]
                & $0.70\pm0.02$ &  (2)\\
\hline

@@

    %  &RA &[degree] &  \phantom{-}84.73 % $\pm$  0.03
   %                     & \multirow{2}{*}{\cite{evans:11}}\\[5pt]
    %  &DEC & [degree] & -69.07 % $\pm$  0.05
     %                   & \\[5pt]
      \multicolumn{5}{l}{\emph{Gaia's relative proper motion for VFTS
      682 }} \\
      \hline
      &$\delta\mu_\mathrm{RA}$  &[$\masyr$] & 0.103$\pm$0.080 & (1,4) \\
      &$\delta\mu_\mathrm{DEC}$  &[$\masyr$] & 0.086$\pm$0.100 &  (1,4) \\
      &$\delta\mu_\mathrm{}$  &[$\masyr$] & \SdM{0.13}$\pm$\SdM{0.09} &  (1,4) \\
                 &$v_\mathrm{2d}$  &[$\kms$] & \SdM{32}$\pm$\SdM{21} & (1,4)\\  
                     &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
                         &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
 %

 \hline     
          \multicolumn{5}{l}{\emph{HST's independent determination of
      the relative proper motion for VFTS 682}} \\
            \hline
      &$\delta\mu_\mathrm{RA, HST}$  &[$\masyr$] & 0.01$\pm$0.13 & (3) \\
      &$\delta\mu_\mathrm{DEC, HST}$  &[$\masyr$] & 0.20$\pm$0.10 &  (3) \\
       &$\delta\mu_\mathrm{HST}$  &[$\masyr$] & 0.20$\pm$0.10 &  (3) \\
                  &$v_\mathrm{2d, HST}$  &[$\kms$] & \SdM{$47\pm24$} & (3)\\  
                      &$\theta^{*}$  &[degrees] &  \SdM{xxx}$\pm$\SdM{xxx}  & (1,4)\\  
                          &$v_\mathrm{2d, \parallel}$  &[$\kms$] & \SdM{xxx}$\pm$\SdM{xxx} & (1,4)\\  
 %

@@

    \end{tabular}
    \tablecomments
    { Proper motions are based on Gaia data unless indicated otherwise in subscript. The angle $\theta$ is defined such that $\theta=0$ corresponds to the direction pointing away from R136.
    (1) \cite{brown:18}, 
    (2) \cite{lennon:18}, 
    (3) \cite{platais:18},  
    (4) {\color{blue}This work}
    %The error on the RA and DEC positions, are of order $\sim$$10^{-2}\masyr$ in \emph{Gaia} DR2. %REPETITION: %  We adopt the peculiar
      % radial velocity calculated using the  NV $\lambda4944$ line, which is
      % less sensitive to the wind velocity structure (see also \Secref{data:vfts683}).
    }
  \end{center}
  \label{tab:vfts682}
\end{table}

@@

The WNh5 star VFTS682, located at right ascension (RA)
05$^\mathrm{h}$38$^\mathrm{m}$55.510$^\mathrm{s}$  and declination
(DEC) \mbox{-69$^\mathrm{o}$04'26.72''} J2000, was observed as part of the multi-epoch, spectroscopic VFTS campaign covering $\lambda$4000--7000 \citep[][]{evans:11}. 
%
%of spectral type WNh5 using multi-epoch spectra covering $\lambda$4000--7000 from the
%VFTS survey. % The spectra available covered $\lambda$4000--7000 and
% were taken at multiple epochs.
%
\citet{bestenlehner:11}  analyzed the spectra to infer the stellar parameters and measure an extinction of $A_V=4.45\pm0.12$, implying a luminosity of $\log_{10}(L/L_\odot) =  6.5\pm0.2$, making this one of the brightest stars in the region. The star is unlikely to have a close companion, because of the absence of clear radial velocity variations \citet{bestenlehner:11}.  Bayesian fits of the stellar parameters against evolutionary tracks \citep{brott:11, kohler:15} using the Bonsaii code \citep{schneider:17, schneider:18} provide estimates for the age, present mass and initial mass, indicating that this star among the most massive stars known, see \Tabref{tab:star} for an overview of the parameters.
